Sampling requirements to detect ecosystem change

### 2020-02-18 R/manuscript_draft.Rmd
added chunk to install packages needed and made suggestions to case study text

---
bibliography: Whitebib.bib
csl: ecology-letters.csl
editor_options:
  chunk_output_type: console
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{float} \usepackage{lineno} \usepackage{setspace}\doublespacing
  \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{,} \usepackage{color} \usepackage{totcount}
  \newtotcounter{citenum} \def\oldcite{} \let\oldcite=\bibcite \def\bibcite{\stepcounter{citenum}\oldcite}
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
---

@@

@@

An application of our study we foresee applies to paleosequences. Sediment cores are taken from aquatic environment to reconstruct past ecosystems, in order to extend our perspective on ecosystem trajectories. Samples can be taken before and after perturbation to compare community compostion (e.g., ), or at continuously to identify the timing of change, and contribution of environmental variable in driving these changes (e.g., @perga_high-resolution_2015).

We tested the three sampling methods on a real community time-series, from Lake Varese (IT). Lake Varese is a small (14.8 km^2^), deep (z$_{max}$= 26m), monomictic lake, in the subalpine region of north-western Italy (238 m asl). It underwent hyper-eutrophication over the 20^th^ century due to increase in nutrient loads from the watershed (ref). Nutrient status was responsible for restructuration of the lake communities across trophic levels (e.g., Crosta 1999, @bruel_seeking_2018). Air temperature is now driving changes in plankton communities (@bruel_seeking_2018).

Cladoceran assemblage was reconstructed at a 2-3 years resolution, from a 74-cm sediment core (VAR10-10) covering the 1816(±26)—2010 period (@bruel_seeking_2018). We defined the "true" changepoint as the one identified from the complete record, in 1983/1988. 
Changepoints were detected on first component computed from the data sub-sampled. We used the same ordination method as in Bruel et al (2018), but note that DCA is best used to conserve multiple gradients. As a result, several changepoints were detected on the first component (in 1926/1928, 1946/1948 and 1983/1988). We conducted the analysis on the second component as well, where only one changepoint was detected (in 1944/1946).

We used the full record (74 observation) and subsampled following the three strategies described earlier (random, regular, iterative), varrying the maximum number analyzed from 5 to 15. Iterative sampling final number could be lower than maximum number if both samples framing the shift were analyzed.

>>>>

---
bibliography: Whitebib.bib
csl: ecology-letters.csl
editor_options:
  chunk_output_type: console
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{float} \usepackage{lineno} \usepackage{setspace}\doublespacing
  \usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{,} \usepackage{color} \usepackage{totcount}
  \newtotcounter{citenum} \def\oldcite{} \let\oldcite=\bibcite \def\bibcite{\stepcounter{citenum}\oldcite} \newcommand\easton[1]{\textcolor{red}{#1}} \newcommand\rosalie[1]{\textcolor{blue}{#1}}
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
---

@@

```{r load_packages,echo=F,warning=F,message=F}
if (!require("pacman")){install.packages("pacman");library(pacman)}

pacman::p_load(patchwork,stickylabeller)
```

@@

\easton{As one application of our approach, we examined paleosequence data.} Sediment cores are taken from aquatic environment to reconstruct past ecosystems, in order to extend our perspective on ecosystem trajectories. Samples can be taken before and after perturbation to compare community compostion (e.g., ), or at continuously to identify the timing of change, and contribution of environmental variable in driving these changes (e.g., @perga_high-resolution_2015).

\easton{In line with the simulation results, we tested different sampling methods} on a real community time-series, from Lake Varese (IT). Lake Varese is a small (14.8 km^2^), deep (z$_{max}$= 26m), monomictic lake, in the subalpine region of north-western Italy (238 m asl). It underwent hyper-eutrophication over the 20^th^ century due to increase in nutrient loads from the watershed (ref). Nutrient status was responsible for restructuration of the lake communities across trophic levels (e.g., Crosta 1999, @bruel_seeking_2018). Air temperature is now driving changes in plankton communities (@bruel_seeking_2018).

Previously, Cladoceran assemblage was reconstructed at a 2-3 years resolution, from a 74-cm sediment core (VAR10-10) covering the 1816(±26)—2010 period [@bruel_seeking_2018]. We detected changepoints on the first component computed from the assemblage data. We used the same ordination method as in Bruel et al (2018), but note that Detrended Correspondence Analysis (DCA) is best used to conserve multiple gradients (\easton{Rosalie: can you explain this a bit more, I'm not sure what it means}). As a result, several changepoints were detected on the first component (in 1926/1928, 1946/1948 and 1983/1988). We defined these as the "true" changepoints given they came from an analysis of the complete data. The 1983-1988 changepoint was the largest in magnitude. (\easton{Rosalie: does this mean that the changepoint occured sometime between 1983 and 1988}). <!--We conducted the analysis on the second component as well, where only one changepoint was detected (in 1944/1946).-->

We used the full record (74 observations) and subsampled following the three strategies described earlier (random, regular, iterative), varrying the maximum number analyzed from 5 to 15. Iterative sampling final number could be lower than maximum number if both samples framing the shift were analyzed.
